Model Recognition as Planning

### 2019-02-27 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: typos by reviewer#2

The exact computation of $P(\mathcal{O}|\mathcal{M})$ with equation (2) is intractable. For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge, infinite in the case of planning problems without dead-ends~\cite{lesh1995sound}. Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables. Instead, our approach is to estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models. {\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, which have been proved successful for {\em pattern recognition}~\cite{MasekP80,Bunke97}. In this work, we assess the cost of {\em transforming} $\mathcal{M}$ into $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.

@@

$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the edition that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$. This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of the model $\mathcal{M}$~\cite{yang2007learning,SohrabiRU16}. Our definition of {\em observation edit distance} is more practical since the size of ${\mathcal I}_{\Psi,\xi}$ is usually much smaller than $F$ (the number of variables in the action schemata should normally be lower than the number of objects in a planning problem).

@@

Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.

@@

In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.

@@

\subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}. Other planners, such as FastDownward were also tested but provided worse experimental results. We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance obtained. The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/model-recognition} so any experimental data reported in the paper can be reproduced.

@@

An interesting research line related to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, in which model edition is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models. {\em Model recognition}, however, conforms every input model $\mathcal{M}$ with another model $\mathcal{M}'$ that is not given as input but instead computed for every $\mathcal{M}$, and which is consistent with a partial observation of a plan execution.

>>>>

The exact computation of $P(\mathcal{O}|\mathcal{M})$ with equation (2) is intractable. For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge, infinite in the case of planning problems without dead-ends~\cite{lesh1995sound}. Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables. Instead, our approach is to estimate $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} defined for \strips\ planning models. {\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, which have been proved successful for {\em pattern recognition}~\cite{MasekP80,Bunke97}. In this work, we assess the cost of {\em transforming} $\mathcal{M}$ into $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.

@@

$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the editing that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$. This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of the model $\mathcal{M}$~\cite{yang2007learning,SohrabiRU16}. Our definition of {\em observation edit distance} is more practical since the size of ${\mathcal I}_{\Psi,\xi}$ is usually much smaller than $F$ (the number of variables in the action schemata should normally be lower than the number of objects in a planning problem).

@@

Now we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.

@@

In other words, we are modeling the editing of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.

@@

\subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}. Other planners, such as FastDownward were also tested but provided worse experimental results. We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance. The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/model-recognition} so any experimental data reported in the paper can be reproduced.

@@

An interesting research line related to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, in which model editing is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models. {\em Model recognition}, however, conforms every input model $\mathcal{M}$ with another model $\mathcal{M}'$ that is not given as input but instead computed for every $\mathcal{M}$, and which is consistent with a partial observation of a plan execution.
